---
title: "Analysis of treadmill fitness test data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysis of treadmill fitness test data}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6.5,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(TreadmillFitnessTest)
```

# How to figure out startminutes and endminutes

Useful in case the run was started before the start of the actual step test,
or ended after.
```{r}
fname = system.file("extdata/2023-01-16.tcx", package = "TreadmillFitnessTest")
tmp <- prepdata(fname, startminutes = 0, endminutes = 100,
          useDeviceSpeed = TRUE)
plot(tmp$minutes, tmp$speed)
plot(tmp$minutes, tmp$cadence_running)
plot(tmp$minutes, tmp$heart_rate)
```

By iteratively adjusting the `startminutes` and `endminutes` and replotting, or
just replotting with adjusted axes, I found that the correct start time was at 
0.15 minutes and correct end time at 15 minutes.

# January 2023 test

Import the data and show the same plots.

```{r}
dat202301 <- prepdata(fname, startminutes = 0.15, endminutes = 15,
          useDeviceSpeed = FALSE)
(dat202301$date = substr(dat202301$time[1], 1, 10))
```

```{r}
plot(dat202301$minutes, dat202301$speed)
plot(dat202301$minutes, dat202301$cadence_running)
plot(dat202301$minutes, dat202301$heart_rate)
```

## Model fitting

This model uses all available data points:
```{r}
fitmodel(dat202301, alldata = TRUE)
```

And this model uses only the final 5 measurements in each step:
```{r}
fitmodel(dat202301, alldata = FALSE)
```

# Sept 15, 2023 test

```{r}
fname = system.file("extdata/2023-09-15.tcx", package = "TreadmillFitnessTest")
dat202309 <- prepdata(fname, startminutes = 23.8, endminutes = 40.1,
          useDeviceSpeed = FALSE)
dat202309$date = substr(dat202309$time[1], 1, 10)
with(dat202309, plot(minutes, speed))
```

## Model fitting

```{r}
fitmodel(dat202309, alldata = TRUE)
fitmodel(dat202309, alldata = FALSE)
```


# 2022-23 together

Some plots demonstrating comparison of two tests. First join the two data.frames,
and convert date and speed to factors (to make plots appear the way I want them to).

```{r}
xall <- full_join(x=dat202309, y=dat202301) |>
    mutate(date = factor(date)) |>
    mutate(speed = factor(speed))
```


## Boxplot of heart rate vs speed

```{r, fig.width=6.5, fig.height=6.3}
ggplot(xall, aes(x=speed, y=heart_rate, fill = date)) +
    geom_boxplot()  +
    scale_y_continuous(breaks=seq(90, 200, by=10))
```
## Scatterplot with Loess smoothing curve for heart rate vs speed

```{r, fig.width=6.5, fig.height=6.3}
ggplot(xall, aes(x = minutes, y = heart_rate, color = date)) +
    geom_point(size = 0.5)  +
    geom_smooth() +
    scale_y_continuous(breaks = seq(90, 200, by = 10), name = "Heart Rate (bpm)") +
    scale_x_continuous(breaks = seq(0, 16.5, by = 1.5), name = "Time (minutes)", 
                       sec.axis = sec_axis( ~ . / 1.5 + 6, 
                                            name = "speed (km/h)", 
                                            breaks = seq(6, 16, by = 1)))
```

## Scatterplot with Loess smoothing curve for running cadence vs speed

```{r, fig.width=6.5, fig.height=6.3}
ggplot(xall, aes(x = minutes, y = 2 * cadence_running, color = date)) +
    geom_point(size = 0.5)  +
    geom_smooth() +
    scale_y_continuous(breaks = seq(150, 200, by = 10), name = "Cadence (spm)") +
    scale_x_continuous(breaks = seq(0, 16.5, by = 1.5), name = "Time (minutes)", 
                       sec.axis = sec_axis( ~ . / 1.5 + 6, 
                                            name = "speed (km/h)", 
                                            breaks = seq(6, 16, by = 1)))
```
